---
title: "Results figures for EDR Appendix"
author: "Lauren Forrow"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(magrittr)
library(openxlsx)
library(ggh4x)
```

## Goal

This script creates figures to include in the EDR appendix/manuscript describing Laura Hatfield's simulations comparing internal and external comparison group designs.  Because the target journal limits the number of figures, we want to be clear and compact.

## Inputs

Read in the final simulation output, along with the file describing each scenario
```{r inputs}
dir_root <- "N:/Project/50886_Primary_Care_First/Restricted/SASFARM/DATA/Impacts/Matching/PCFG/Simulations"
sim_res <- read.csv(file.path(dir_root, "Data", "2021-09-20_simulation_summaries.csv"))
sim_pars <- read.xlsx(file.path(dir_root, "Data", "2021-09-19_matching_scenarios_params.xlsx"), sheet = "Sheet1")
```

## Processing

Create bias squared and variance elements as components of MSE, and check against actual RMSE.
```{r}
sim_res_wide <- sim_res %>%
  pivot_wider(names_from = "statistic", values_from = "value") %>%
  mutate(bias_sq = `abs.mean.bias`^2,
         var = sd^2,
         check = rmse - sqrt(bias_sq + var),
         group_lab = factor(ifelse(controls == "External", "Out-of-market", "Within-market"),
                            levels = c("Within-market", "Out-of-market")))

summary(sim_res_wide$check)
```

That's pretty close.  Reformat long.
```{r}
sim_res_long <- sim_res_wide %>%
  select(-check) %>%
  pivot_longer(mean.abs.bias:var, names_to = "statistic", values_to = "value")
```


Merge the parameters information onto the results.
```{r}
pars_wider <- sim_pars %>%
  select(-X4)%>%
  mutate(Param = gsub("\\.", "", Param)) %>% #drop punctuation to make easier colnames
  pivot_wider(names_from = "Param", values_from = "Value")

full_res <- sim_res_long %>%
  left_join(pars_wider, by = "Scenario")
  
```

Clean up labels for plotting.
```{r}
full_res %<>%
  mutate(stat_lab = factor(case_when(statistic == "mean.abs.bias"      ~ "MAB",
                                     statistic == "abs.mean.bias"      ~ "AMB",
                                     statistic == "mean.bias"          ~ "Mean bias",
                                     statistic == "sd"                 ~ "Standard deviation",
                                     statistic == "rmse"               ~ "RMSE",
                                     statistic == "bias_sq"            ~ "Squared bias", 
                                     statistic == "var"                ~ "Variance"), 
                           levels = c("Mean bias", "MAB", "AMB", "Standard deviation", "RMSE", 
                                      "Variance", "Squared bias")),
         year_lab = paste("Year", year),
         cov_lab = factor(ifelse(covid == "covid", "With COVID-19", "Without COVID-19"),
                          levels = c("Without COVID-19", "With COVID-19")))
```


## Plots

### Main text

For the main text of the EDR appendix/manuscript, we want to show just a small number of scenarios that illustrate our main point.  For the appendix, we want to show more of a range.

Compact version.
```{r, fig.width = 8, fig.height = 5}
full_res %>% 
  filter(statistic %in% c("bias_sq", "var"), 
         year %in% c(1, 5),
         Scenario %in% c("C.1.1", "B.3.1")) %>%  #drop D.1 break-even
  # mutate(scen_num = as.numeric(as.factor(Scenario)),
  #        scen_lab = paste("Scenario", scen_num))%>%
  mutate(scen_lab = factor(case_when(Scenario == "C.1.1"     ~ "Moderate participation bias",
                              Scenario == "B.3.1"     ~ "Equal slope differences"),
                           levels = c("Equal slope differences", "Moderate participation bias"))) %>%
  ggplot(., aes(x = group_lab, y = value, fill = stat_lab)) +
  theme_bw() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_col(color = "black") +
  facet_nested(cov_lab~scen_lab + year_lab) +
  ylab("Mean squared error") +
  xlab("Comparison group") +
  scale_fill_manual(name = "Source of error", values = c("#0b2949", "#d9e8f9"), breaks = c("Squared bias", "Variance"))

ggsave(file.path(dir_root, "Output", "results_fig_compact3_mprcolors.png"), height = 5, width = 8, units = 'in', dpi = 450)
```

Fuller version.
```{r, fig.width = 9, fig.height = 6}
full_res %>% 
  filter(statistic %in% c("mean.abs.bias", "rmse"), 
         year %in% c(1, 5),
         Scenario %in% c("B.2.1", "B.3.1", "D.1")) %>%
  # mutate(scen_num = as.numeric(as.factor(Scenario)),
  #        scen_lab = paste("Scenario", scen_num))%>%
  mutate(scen_lab = factor(case_when(Scenario == "B.1.1"     ~ "Parallel out-of-market",
                              Scenario == "B.3.1"     ~ "Equal slope differences",
                              Scenario == "D.1"       ~ "Break-even"),
                           levels = c("Parallel out-of-market", "Equal slope differences", "Break-even"))) %>%
  ggplot(., aes(x = stat_lab, y = value, fill = group_lab)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  geom_col(position = position_dodge(width = 0.9)) +
  facet_nested(cov_lab~scen_lab + year_lab) +
  ylab(NULL) +
  xlab(NULL) +
  scale_fill_manual(name = "Comparison group", values = c("gray33", "gray67"))

ggsave(file.path(dir_root, "Output", "results_fig_expanded2.png"), dpi = 450)
```

### For Appendix

In the appendix, let's show almost all the scenarios, excluding only those that are duplicative.  Split up by scenarios where gamma is negative vs. positive, for those that have both.

```{r, fig.width = 8, fig.height = 11}

# create scenario labels that are consistent between figures

scen_map <- full_res %>% 
  filter(!(Scenario %in% c("D.4", "D.5", "D.7"))) %>%  #drop a few duplicative ones
  mutate(scen_num = as.numeric(as.factor(Scenario)),
         scen_lab = factor(paste("Scenario", scen_num),
                           levels = paste("Scenario", 1:max(scen_num)))) %>%
  distinct(scen_num, scen_lab, Scenario, gamma2, theta2)%>%
  select(scen_num, scen_lab, Scenario, gamma2, theta2)

scen_map %>%
  #put parameters on annual scale
  mutate(gamma2 = 4*gamma2,
         theta2 = 4*theta2)%>%
  write.csv(., file = file.path(dir_root, "Output", "scenario_numbers.csv"), row.names = FALSE)

```

Scenarios with small participation bias
```{r, fig.height = 10, fig.width = 8}
full_res %>% 
  filter(statistic %in% c("bias_sq", "var"), 
         year %in% c(1, 5),
         !(Scenario %in% c("D.4", "D.5", "D.7"))) %>%  #drop a few duplicative ones
  left_join(scen_map, by = "Scenario") %>%
  filter(scen_num %in% 1:6)%>%
  ggplot(., aes(x = group_lab, y = value, fill = stat_lab)) +
  theme_bw() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_col(color = "black") +
  facet_nested(scen_lab ~ cov_lab + year_lab) +
  ylab("Mean squared error") +
  xlab("Comparison group") +
  scale_fill_manual(name = "Source of error", values = c("#0b2949", "#d9e8f9"), breaks = c("Squared bias", "Variance"))

ggsave(file.path(dir_root, "Output", "results_fig_all_small_mprcolors.png"), height = 10, width = 8, units = 'in', dpi = 450)
```

Scenarios with moderate participation bias
```{r, fig.height = 10, fig.width = 8}
full_res %>% 
  filter(statistic %in% c("bias_sq", "var"), 
         year %in% c(1, 5),
         !(Scenario %in% c("D.4", "D.5", "D.7"))) %>%  #drop a few duplicative ones
  left_join(scen_map, by = "Scenario") %>%
  filter(scen_num %in% 7:12)%>%
  ggplot(., aes(x = group_lab, y = value, fill = stat_lab)) +
  theme_bw() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_col(color = "black") +
  facet_nested(scen_lab ~ cov_lab + year_lab) +
  ylab("Mean squared error") +
  xlab("Comparison group") +
  scale_fill_manual(name = "Source of error", values = c("#0b2949", "#d9e8f9"), breaks = c("Squared bias", "Variance"))

ggsave(file.path(dir_root, "Output", "results_fig_all_mod_mprcolors.png"), height = 10, width = 8, units = 'in', dpi = 450)
```

"Break-even" scenarios
```{r, fig.height = 10, fig.width = 8}
full_res %>% 
  filter(statistic %in% c("bias_sq", "var"), 
         year %in% c(1, 5),
         !(Scenario %in% c("D.4", "D.5", "D.7"))) %>%  #drop a few duplicative ones
  left_join(scen_map, by = "Scenario") %>%
  filter(scen_num %in% 13:max(scen_num))%>%
  ggplot(., aes(x = group_lab, y = value, fill = stat_lab)) +
  theme_bw() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_col(color = "black") +
  facet_nested(scen_lab ~ cov_lab + year_lab) +
  ylab("Mean squared error") +
  xlab("Comparison group") +
  scale_fill_manual(name = "Source of error", values = c("#0b2949", "#d9e8f9"), breaks = c("Squared bias", "Variance"))

ggsave(file.path(dir_root, "Output", "results_fig_all_breakeven_mprcolors.png"), height = 8, width = 8, units = 'in', dpi = 450)
```

